\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-02-05 tex/overview.tex
minor edits

The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be further abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc). This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations. These three symmetries represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated \reffigscc. Two of these three are also symmetries with respect to dynamical robustness. \ref{fig:robustnesssymmetries} shows an example of these symmetries applied to a specific interaction graph.

@@

We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$. The entries of the Jacobian are sampled independently from a generic probability distribution. Examples of vector fields that may result from this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}. The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
\begin{equation}\label{eq:robschematic}
R_{tot} = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
\end{equation}
For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.  From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.  The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.  Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.

@@

This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size. This analytical result predicts that any network whose associated dynamical system has the interaction graph \reffigscc $\,$ top will be more robust than those associated to any of the other interaction graphs in \reffigscc. Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system. The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.

>>>>

The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc). This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations. These three symmetries represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated \reffigscc. Two of these three are also symmetries with respect to dynamical robustness. \ref{fig:robustnesssymmetries} shows an example of these symmetries applied to a specific interaction graph.

@@

We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$. The entries of the Jacobian are sampled independently from a generic probability distribution. Examples of vector fields that may result from this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}. The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
\begin{equation}\label{eq:robschematic}
R_{tot} = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
\end{equation}
For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.  From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.  The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.  Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc{}, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.

@@

This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size. This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the reflexive total ordering like that of \reffigscc{} top will be more robust than those associated to any of the other interaction graphs in \reffigscc{}. Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system. The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
